none

### 2024-07-05 paper/neural_2024/poster_text/intro.tex
Adds poster for Neural Computation 2024

The axons of retinal ganglion cells form an ordered map of the retina
in the optic tectum. Surgical and genetic experiments have revealed
several key mechanisms that contribute to retinotectal map
development, and computational modelling studies have demonstrated how
maps can be shaped by their interactions. Models derived from Sperry's
classic chemoaffinity theory can reproduce retinotopy, and Gierer
later showed how an extension to include competition between growing
axons could also account for a subset of the disordered maps that have
resulted from surgical manipulation. Further extensions have each
accounted for additional datapoints, but none has been able to explain
the full range of maps observed experimentally. Here we present a
unified model of retinotectal development, assuming only local
interactions between growing axons, which can account for the full
range of maps resulting from surgical manipulation. This model extends
the `potential functions' of Gierer's original to describe
self-organisation across a 2D cortical sheet, and it represents a
coordination of axonal growth by orthogonal pairs of opposing ligand
gradients in the tectum and corresponding receptor gradients in the
retina. In addition, we present progress towards accounting for the
full range of maps that have resulted from genetic manipulation, by
incorporating a mechanistic description of receptor ligand
interactions.

%\bibliography{RetinoTectal}

>>>>

%\section{Introduction}

\subsection{A summary of this work}

\begin{itemize}
\item Retinal cells grow axons that form an ordered map in the optic tectum
\item Sperry, Gierer, Prestige, Willshaw and many others have all sought to model this
  developmental process to better understand how brains wire themselves up
\item Surgical and genetic manipulation experiments have provided a wide range
  of developmental responses that these models should be able to describe
\item We want to make a model based on local interactions that can account for
  all of the experimental results
\end{itemize}

\subsection{A brief history of retinotectal studies}

(For reviews see \citealp{Prestige1975,Goodhill2005})

\begin{itemize}
\item 1940s: Sperry's surgical experimental and theoretical work
  \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}
\item 1970s: Marker-induction models \citep{Prestige1975,Willshaw2006}
\item 1980s: Chemotaxis models \citep{Gierer1983,Gierer1987}, branch arrow
  models \citep{overton_extended_1982}
\item 1990s: Discovery of the ephrin signalling molecules \citep{nakamoto_topographically_1996}
\item 2000s: Important genetic manipulation studies
  \citep{brown_topographic_2000,reber_relative_2004}; Hebbian activity models \citep{Tsigankov2006,Tsigankov2010}
\item 2010s: Competition models \citep{Triplett2011,Simpson2011}; comparative
  studies \citep{hjorth_quantitative_2015}
\item Recently: Re-evaluation of chemoaffinity \citep{Sterratt2013,naoki_revisiting_2017}
\end{itemize}

One of the more abstract models \citep{Triplett2011} was shown by
\cite{hjorth_quantitative_2015} to be the only one able to explain the `EphA3
collapse point'. However, although the model is evaluated numerically (by
iterative switching), this procedure is not necessarily an explicit
representation of the process of development \citep{Wilson2015}. We wanted to
find a mechanistic model which can be directly related to the behaviour of
developing axon processes. Starting with a modified version of \cite{James2020},
and influenced by the \cite{Simpson2011}
we built a Gierer-like gradient climbing model in which we attempted to
express all the interactions of the axon branches in terms of receptor-ligand
signalling.

\bibliography{RetinoTectal}

### 2024-07-05 paper/neural_2024/poster_text/results2.tex
Adds poster for Neural Computation 2024

Each of the $M$ RGC growth cones carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
%
We defined 4 receptor expression gradients (each varying only wrt one spatial dimension) arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$. $r_2$, whose gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,1B). Thus the model is of \emph{differential chemotaxis}.

Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in orthogonal pairs of gradients.
Although several studies model tectal ligand expression with exponential
functions \citep{koulakov_stochastic_2004}, the experimental evidence for the
form of tectal ligand expression is more ambiguous than that for retinal
receptor expression (For example data, see Fig.\,9 which shows expression patterns for EphB/ephrinB).
We kept in mind the possibility that tectal ligand expression may be modelled
by another function, but initially, we set it to the same exponential used for
retinal expression and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in
place of $u$ to obtain the ligand expression functions for $L_0$, $L_1$, $L_2$
and $L_3$ (Fig.\,1).

>>>>

Each of the $M$ RGC growth cones carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
%
We defined 4 receptor expression gradients arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$. $r_2$ was orthogonal to $r_3$ (Fig.\,1B). Thus the model is of \emph{differential chemotaxis}.

%Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in
%orthogonal pairs of gradients.  Several studies model tectal ligand expression
%with exponential functions \citep{koulakov_stochastic_2004}. Although the
%experimental evidence to support this is lean, we also set $L_i$ to the same
%exponential used for retinal expression.
